---
title: "Homework 02 ReadMe"
output:
  github_document: default
  keep_md: default
---

# Doing Homework 02

## Installing gapminder and tidyverse.

```{r}
install.packages("gapminder")
library(gapminder)
```


```{r}
install.packages("tidyverse")
library(tidyverse)
```

## (A) Exploring attributes of gapminder

### (1) Type of Gapminder
```{r}
typeof(gapminder)
```

### (2) Class of Gapminder
```{r}
class(gapminder)
```

### (3) Variables / Columns
```{r}
ncol(gapminder)
```

### (4) Rows / Observations
```{r}
nrow(gapminder)
```

### (5) Alternative ways of determining "size" or "extent"
```{r}
dim(gapminder)
```

```{r}
summary(gapminder)
```

### (6) Type of each variable?
```{r}
sapply(gapminder, typeof)
```

#### Reflections of the Process
- This section took me awhile since I have forgotten almost all of these functions and had to go back to the notes. As you might have guessed, I am new at R. 
- I also consulted google a lot. My seatmate, Cindy, who was the one of the peer reviewers of my first homework assignment, also helped a lot. Shoutout to Cindy! I also learned from the two classmates whose first homework I reviewed.
- My first homework was done haphazardly. I will eventually edit the main ReadMe file. I hope this homework is easier to follow for the reviewers and markers.


! https://en.wikipedia.org/wiki/Keep_Calm_and_Carry_On#/media/File:Keep-calm-and-carry-on-scan.jpg

## (B) Exploring individual variables of gapminder

### (1) Possible values (or range, whichever is appropriate) of each variable?
```{r}
max <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),max)
min <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),min)
#Range of Years
c(min[1],max[1])
#Range of lifeExp
c(min[1],max[1])
#Range of pop
c(min[1],max[1])
#Range of gdpPercap
c(min[1],max[1])
```


``` r


```

    ## year year 
    ## 2007 1952

``` r
#Range of LifeExp
c(max[2],min[2])
```

    ## lifeExp lifeExp 
    ##  82.603  23.599

``` r
#Range of Pop
c(max[3],min[3])
```

    ##        pop        pop 
    ## 1318683096      60011

``` r
#Range of gdpPercap
c(max[4],min[4])
```

    ##   gdpPercap   gdpPercap 
    ## 113523.1329    241.1659

``` r
table(gapminder$country)
```

    ## 
    ##              Afghanistan                  Albania                  Algeria 
    ##                       12                       12                       12 
    ##                   Angola                Argentina                Australia 
    ##                       12                       12                       12 
    ##                  Austria                  Bahrain               Bangladesh 
    ##                       12                       12                       12 
    ##                  Belgium                    Benin                  Bolivia 
    ##                       12                       12                       12 
    ##   Bosnia and Herzegovina                 Botswana                   Brazil 
    ##                       12                       12                       12 
    ##                 Bulgaria             Burkina Faso                  Burundi 
    ##                       12                       12                       12 
    ##                 Cambodia                 Cameroon                   Canada 
    ##                       12                       12                       12 
    ## Central African Republic                     Chad                    Chile 
    ##                       12                       12                       12 
    ##                    China                 Colombia                  Comoros 
    ##                       12                       12                       12 
    ##         Congo, Dem. Rep.              Congo, Rep.               Costa Rica 
    ##                       12                       12                       12 
    ##            Cote d'Ivoire                  Croatia                     Cuba 
    ##                       12                       12                       12 
    ##           Czech Republic                  Denmark                 Djibouti 
    ##                       12                       12                       12 
    ##       Dominican Republic                  Ecuador                    Egypt 
    ##                       12                       12                       12 
    ##              El Salvador        Equatorial Guinea                  Eritrea 
    ##                       12                       12                       12 
    ##                 Ethiopia                  Finland                   France 
    ##                       12                       12                       12 
    ##                    Gabon                   Gambia                  Germany 
    ##                       12                       12                       12 
    ##                    Ghana                   Greece                Guatemala 
    ##                       12                       12                       12 
    ##                   Guinea            Guinea-Bissau                    Haiti 
    ##                       12                       12                       12 
    ##                 Honduras         Hong Kong, China                  Hungary 
    ##                       12                       12                       12 
    ##                  Iceland                    India                Indonesia 
    ##                       12                       12                       12 
    ##                     Iran                     Iraq                  Ireland 
    ##                       12                       12                       12 
    ##                   Israel                    Italy                  Jamaica 
    ##                       12                       12                       12 
    ##                    Japan                   Jordan                    Kenya 
    ##                       12                       12                       12 
    ##         Korea, Dem. Rep.              Korea, Rep.                   Kuwait 
    ##                       12                       12                       12 
    ##                  Lebanon                  Lesotho                  Liberia 
    ##                       12                       12                       12 
    ##                    Libya               Madagascar                   Malawi 
    ##                       12                       12                       12 
    ##                 Malaysia                     Mali               Mauritania 
    ##                       12                       12                       12 
    ##                Mauritius                   Mexico                 Mongolia 
    ##                       12                       12                       12 
    ##               Montenegro                  Morocco               Mozambique 
    ##                       12                       12                       12 
    ##                  Myanmar                  Namibia                    Nepal 
    ##                       12                       12                       12 
    ##              Netherlands              New Zealand                Nicaragua 
    ##                       12                       12                       12 
    ##                    Niger                  Nigeria                   Norway 
    ##                       12                       12                       12 
    ##                     Oman                 Pakistan                   Panama 
    ##                       12                       12                       12 
    ##                 Paraguay                     Peru              Philippines 
    ##                       12                       12                       12 
    ##                   Poland                 Portugal              Puerto Rico 
    ##                       12                       12                       12 
    ##                  Reunion                  Romania                   Rwanda 
    ##                       12                       12                       12 
    ##    Sao Tome and Principe             Saudi Arabia                  Senegal 
    ##                       12                       12                       12 
    ##                   Serbia             Sierra Leone                Singapore 
    ##                       12                       12                       12 
    ##          Slovak Republic                 Slovenia                  Somalia 
    ##                       12                       12                       12 
    ##             South Africa                    Spain                Sri Lanka 
    ##                       12                       12                       12 
    ##                    Sudan                Swaziland                   Sweden 
    ##                       12                       12                       12 
    ##              Switzerland                    Syria                   Taiwan 
    ##                       12                       12                       12 
    ##                 Tanzania                 Thailand                     Togo 
    ##                       12                       12                       12 
    ##      Trinidad and Tobago                  Tunisia                   Turkey 
    ##                       12                       12                       12 
    ##                   Uganda           United Kingdom            United States 
    ##                       12                       12                       12 
    ##                  Uruguay                Venezuela                  Vietnam 
    ##                       12                       12                       12 
    ##       West Bank and Gaza              Yemen, Rep.                   Zambia 
    ##                       12                       12                       12 
    ##                 Zimbabwe 
    ##                       12

``` r
table(gapminder$continent)
```

    ## 
    ##   Africa Americas     Asia   Europe  Oceania 
    ##      624      300      396      360       24


### (2) Typical values: spread and distribution
```{r}
HW02Plot1 <- ggplot(gapminder, aes(x=continent, y=pop)) + geom_line()
HW02Plot1
```

```{r}
HW02Plot2 <- ggplot(gapminder, aes(x=country, y=gdpPercap)) + geom_point()
HW02Plot2
```

```{r}
HW02Plot3 <- ggplot(gapminder, aes(x=pop, color=continent)) + 
  geom_density()
HW02Plot3
```

#### Reflections on the process
- This is fun. I could do more of this. I am itching to work on my data now and produce these visualizations.

## (C) Exploring various plot types

### (1) Scatterplot of two quantitative variables

```{r}
gapminder %>%
  filter (country %in% c("Philippines", "Canada","South Africa","Peru")) %>% 
  ggplot(aes(x = pop, y = lifeExp)) +
  geom_point(aes(color = country))
```

### (2) A plot of one quantitative variable
```{r}
gapminder %>%
  filter (country %in% c("Philippines", "Canada","South Africa","Peru")) %>% 
  ggplot(aes(x = year, y = pop)) +
  geom_line(aes(color = country))
```

### (3) A plot of one quantitative variable and one categorical variable.
```{r}
gapminder %>%
  filter (country %in% c("Philippines", "Canada","South Africa","Peru")) %>% 
  ggplot(aes(x = year, y = pop, color = country)) +
  geom_boxplot(aes(shape = country))
```

### (4) Using select and filter and piping

```{r}
select(gapminder, country, continent)
```

```{r}
filter(gapminder, 
       continent != "Asia")
```

```{r}
gapminder %>% 
  filter (country %in% c("Philippines", "Peru"),
          year<= 1969, year >= 1960) %>%
  select (country, year, pop)
```

#### Reflections of the Process
- I want to explore more but I don't have time. This has been a very useful homework.
- I am having difficulties pushing this back to GitHub.