---
title: "Hw_03"
output: github_document
---

# Doing Homework 03 and Reflecting on the Process

Loading the libraries
```{r}
library(gapminder)
library(tidyverse)
```


## Task One:
Get the maximum and minimum of GDP per capita for all continents

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(max_gdpPercap = max(gdpPercap))
```
### Note 1:
I reworked a code from cm008 notes. 

```{r}
table(max_gdppercap)
```
### Note 2:
I don't remember this function being discussed in class. So I looked up a code on google. After a few tries, I think I have managed to do a table.

```{r}
knitr::kable(max_gdppercap)
```
### Note 3:
I used the code suggested in the hw03 page substituting my_df to max_gdppercap.

```{r}
maxplot <- ggplot(max_gdppercap,aes(continent, max_gdppercap)) + geom_point()
maxplot
```
### Note 4:
I used geom_point to visualize the difference between max_gdppercap among continents.

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(min_gdpPercap = min(gdpPercap))
table(min_gdppercap)
knitr::kable(min_gdppercap)
ggplot(min_gdppercap,aes(x=continent,y=min_gdPercap)) + geom_point()
```
### Note 5:
For the min_gdpPercap, I copied the codes I used above, changed max to min, and laid out all the codes in one chunk. 

## Task Two: 
How is life expectancy changing over time on different continents?

```{r}
weighted_LifExp <- gapminder %>% 
  group_by(year, continent) %>%
  summarize(weighted_lifExp = weighted.mean(lifeExp,pop))
weighted_LifExp
```
### Note 6:
This took me quite a while to figure out. I had to search for similar codes online that will allow me to come up with this table. 

```{r}
ggplot(weighted_LifExp, aes(year, weighted_lifExp, color=continent)) + geom_line ()
```
### Note 7:
After I figured out how to present weighted_lifeExp in a table, visualizing it came easily. 
  
